Tina Paper In Development

### 2024-10-30 docs/writing/outline.tex
Starting to write introduction and updating outline as I write and reword

\1 Abstract
	\item Flatbed scanner alternative with improved resolution to digitizing tree cookies and cores
	\item Digital twins of samples is important for building data repositories and promoting replicable science
	\item Digitizing wood samples has limits to sample dimensions or lacks affordable implementations
	\item Current alternatives are very expensive (\$70,000USD), lack a high level of detail (printer scanners), or cannot scan cookies (CaptuRING)
	\item Tina is open source and able to be assembled with 3D printed parts, OpenBuilds parts kits, and hand tools % Can we add how much ours costs?
\item Introduction
	\item Tree ring science overview
		\item tree rings provide critical data/insights across multiple fields % Guibal and Guiot
		\item tree rings can help reconstruct past climates (dendroclimatology) % Sheppard
		\item understanding tree growth mechanisms (plant bio/ecology) % Fritts and Swetnam 
	% Consider adding this to below points (Development of scanning alternatives) as one paragraph on its own
	\item Getting these data means measuring tree rings -- quick history
		\item originally done on scopes
		\item images important for several decades now (WinDendro?) 
		\item But new methods coming up often to take advantage of better imaging technology ... 
			\item vessel counting requires detail at the sub-vessel level % detail is only as good as sample prep
		% \item Could add here one sentence on the three methods you mention below depending on what you want to say ... 
	\item Development of scanning alternatives 
		\item beginning with ATRICS
		\item expensive Gigapixel
		\item DIY CaptuRING

@@

@@

		\item allow a method to digitize cookies to have higher quality data than cores alone
		\item avoid size limitations to digitize tree cookies and cores 

@@

>>>>

\1 Abstract
\begin{enumerate}
	\item Flatbed scanner alternative with improved resolution to digitizing tree cookies and cores
	\item Digital twins of samples is important for building data repositories and promoting replicable science
	\item Digitizing wood samples has limits to sample dimensions or lacks affordable implementations
	\item Current alternatives are very expensive (\$70,000USD), lack a high level of detail (printer scanners), or cannot scan cookies (CaptuRING)
	\item Tina is open source and able to be assembled with 3D printed parts, OpenBuilds parts kits, and hand tools % Can we add how much ours costs?
\end{enumerate}
\item Introduction
	\begin{enumerate}
	\item Tree ring science overview
		\begin{enumerate}
		\item tree rings provide critical data/insights across multiple fields % Guibal and Guiot
		\item tree rings can help reconstruct past climates (dendroclimatology) % Sheppard
		\item understanding tree growth mechanisms (plant bio/ecology) % Fritts and Swetnam 
		\end{enumerate}
	% Consider adding this to below points (Development of scanning alternatives) as one paragraph on its own
	\item Getting these data means measuring tree rings -- quick history
		\begin{enumerate}
		\item originally done on scopes (stage micrometer) % Robinson and Evans, https://conservancy.umn.edu/server/api/core/bitstreams/30f91f03-c27b-44de-8659-5156220ce79b/content
		\item eventually there was a shift to analyzing digital images of samples instead of stage micrometers % Guay et al. 1992 
		\item digital copies inherently allow for contributing to data pools  % Rademacher​ et al. WIAD; also International Tree-Ring Data Bank (IRTB) https://www.ncei.noaa.gov/products/paleoclimatology/tree-ring
		\item Development of image acquisition in dendrochronology 
			\begin{enumerate}
			\item two models: flatbed scanner or overlapping images and stitch
			\item beginning with flatbed scanners % WinDendro
				\subitem modern flatbed scanners can achieve high resolution (4200 dpi) but size of samples is limited % see CaptuRING paper for flatbed specs
			\item overlapping images and stitching 
				\subitem ATRICS, Gigapixel, CaptuRING 
				\subitem takes advantage of advanced camera technology
			\end{enumerate}
		% \item Could add here one sentence on the three methods you mention below depending on what you want to say ... 
		\end{enumerate}

@@

		\item extend the capabilities of CaptuRING by allowing to stitch cookies and multiple samples in a queue

@@

		\item increase the maximum size of a sample 

@@

		\item vessel counting requires detail at the sub-vessel level % https://roxas.wsl.ch/en/detail is only as good as sample prep

### 2024-10-30 docs/writing/paper.tex
Starting to write introduction and updating outline as I write and reword

\section{Introduction}
This is the introduction section where you introduce the topic and cite some papers.

>>>>

\section{Introduction}
The study of tree rings has proven useful across multiple fields, proving to be a reliable subject for reconstructing past climates of regional and local environments, as well as a mechanism to understand tree growth response \citep{fritts_dendroclimatology_1971} \citep{williams_using_2010} \citep{guibal_dendrochronology_2021} \citep{sheppard_dendroclimatology_2010}.
To obtain these data from the tree rings, it is necessary to measure tree rings from a tree cookie or core. The first high precision tool for this purpose was a stage micrometer, involving a trained technician to incrementally shift a tree core under the objective of a microscope - informing a computer when a new ring is encountered \citep{robinson_microcomputer_nodate}.
While this method has very high precision, the data is only as accurate as the experience and knowledge of the technician at the time of recording \citep{levanic_atrics_2007}.
The desire to remove repetition of errors in sampling and sampling bias across to individual technicians led researchers to an alternative - image analysis. 
\par
The first step in measuring tree ring width from images requires the digitization of the sample. There are currently two competing models to digitize a sample. The first model is to use a flatbed scanner or similar tool, providing the benefit to

The first approach involved flatbed scanning technology which digitized the surface of a prepared sample \citep{guay_new_1992}. 
Flatbed scanners eventually encountered a resolution insufficiency when analyzing very narrow tree rings on the scale of a few cells. A new method introduced with the ATRICS system increased the resolution of samples by stitching together multiple images from a digital microscope camera \citep{levanic_atrics_2007}. 
ATRICS set an example of the process which has been implemented in two other image acquisition tools, Gigapixel and CaptuRING \citep{griffin_gigapixel_2021, garcia-hidalgo_capturing_2022}.
Both of these tools use professional cameras to take individual images of cores, although CaptuRING requires the technician to transfer the images to a second device and use the proprietary image stitching software PTGUI. 
A budget friendly image acquisition tool with integrated image stitching for both cores and cookies has yet to be released.

% Potentially should have a section on tree ring width measurement software
% The images were then fed to software such as WinDENDRO to have technicians click tree rings on a graphical user interface (GUI) \citep{guay_new_1992}.
% More modern implementations of measurement software now exists which includes width measurement, cross dating, and autodetection of rings as well \citep{maxwell_measuring_2021}.
